---
output: pdf_document
title: Stastistical Inference Course Project
author: Sriram Vadlamani
---

In this project we will investigate the exponential distribution in R and compare it with the 
Central Limit Theorem. The exponential distribution can be expDistributionulated in R with rexp(n, lambda) 
where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation 
is also 1/lambda. Set lambda = 0.2 for all of the simulations.

We will investigate the distribution of averages of 40 exponentials. Lets do the 1000 simulations. 


```{r}
#Load the libraries
library(ggplot2)

#set the seed for reproducability
set.seed(31233)

#set-up the simulaiton
lambda <- 0.2
num_sims <- 1000
sample_size <- 40
expDistribution <- matrix(rexp(num_sims*sample_size, rate=lambda), num_sims, sample_size)
means <- apply(expDistribution, 1, mean)

```

# 1. Sample Mean vs Theoretical Mean

```{r echo=FALSE}

hist(means, breaks=50, prob=TRUE,xlab="")
lines(density(means))
abline(v=1/lambda, col="red") # theoretical center of distribution
xfit <- seq(min(means), max(means), length=100)
yfit <- dnorm(xfit, mean=1/lambda, sd=(1/lambda/sqrt(sample_size)))
lines(xfit, yfit, pch=22, col="red", lty=2)
legend('topright', c("Simulated", "Theoretical"), lty=c(1,2), col=c("black", "red"))
```

The distribution of sample means is centered at `r mean(means)`
and the theoretical center of the distribution is `r 1/lambda`.

# 2. Sample Variance vs Theoretical Variance

The variance of sample means is `r var(means)` where the theoretical variance
of the distribution is $\sigma^2 / n = 1/(\lambda^2 n) = 1/(0.04 \times 40)$ =
`r 1/(0.04 * 40)`.

As per the central limit theorem, the averages of samples follow normal
distribution. The figure above also shows the density computed using the histogram and the
normal density plotted with theoretical mean and variance values. Also, the
q-q plot below suggests the normality.

```{r echo=FALSE}
qqnorm(means)
qqline(means)

```

# 3. Distribution
Let's evaluate the coverage of the confidence interval.
  
  ```{r echo=FALSE}
lambda_vals <- seq(4, 6, by=0.01)
coverage <- sapply(lambda_vals, function(lamb) {
  mu_hats <- rowMeans(matrix(rexp(sample_size*num_sims, rate=0.2),
                             num_sims, sample_size))
  ll <- mu_hats - qnorm(0.975) * sqrt(1/lambda**2/sample_size)
  ul <- mu_hats + qnorm(0.975) * sqrt(1/lambda**2/sample_size)
  mean(ll < lamb & ul > lamb)
})


qplot(lambda_vals, coverage) + geom_hline(yintercept=0.95)
```
As you can see from the plot above, for lambda around 5,
the average of the sample mean falls within the confidence interval, at least 95% of the time.
The true lambda is 5.